Stack Safety as a Security Property

### 2020-07-06 main.tex
Description of eager testing property

\leo{Which of the machine, machine-observation, machine-policy,
  and machine-policy-observation traces should we distinguish
  in metavariables?}

@@

\paragraph*{Testing the conservative policy}

This conservative policy enforces our stack safety policy. Moreover, it enforces
a provably stronger \emph{testing property}.

The properties we give are indeed enforceable.

What is the strongest property enforced by the conservative policy?
%
\rb{Not too much focus on this}

\rb{We have not proved that this policy guarantees our property, but we have
tested it against a stronger property.}

>>>>

\leo{Which of the machine, machine-observation, machine-policy,
  and machine-policy-observation traces should we distinguish
  in metavariables?}
\rb{Should we have notation for traces somewhere here?}

@@

%% Because the access rights of callers subsume those of callees, it is essential
%% that the micropolicy change tags when the property thinks that a call occurs,
%% but it is safe for the micropolicy to fail to reset to the caller's tag when the
%% property thinks a return occurs. (Dually, it is safe for the micropolicy to
%% change tags even without executing a call point, but not for it to reset tags to
%% the caller when the property thinks execution is still in the caller).

%% Because the property does not identify return sites, we have no way to require
%% that ``intended returns'' are tagged correspondingly. Some unintended results
%% might also be so tagged, but as long as the jump has the effect of a legal
%% return, it is OK.

\paragraph*{Testing the conservative policy}
%
This conservative policy is can be shown to satisfy an eager \emph{testing
property}, which enforces lockstep and performs integrity and confidentiality
tests for a stack of running variants at each step.

\definition{Eager integrity test}
  Given two machines $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{eager integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\contour) = \mach'(\contour)
    \]

\definition{Eager confidentiality test}
  Given two machine states $\mach_1$ and $\mach_2$ with stepping states and
  observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$ these satisfy the \emph{eager confidentiality
  test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \obs_1 ~ \mach_1' ~ \mach_2 ~ \obs_2 ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1' = \mach_2'
  \]

\rb{Fix notations, including for properties below.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\definition{Variant stacks}
  The testing property keeps track of execution information on a stack whose
  elements record an initial machine, an initial and current variant of the
  machine, a contour and a return map:
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS (\MACHS \rightarrow \mathcal{P})$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
  Given the current machine state $\mach$, a stack $\vs$ is \emph{well-formed}
  with respect to it if:
  \begin{itemize}
  \item The current variant is indeed a variant of the current state.
  \item The stack is nonempty and the bottom return map is the false property.
  \item The current state is in the trace of every initial machine.
  \item The variant state is in the trace of its own initial variant.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item All contours have public registers.
  \end{itemize}

\definition{Eager testing property}
  Let $\callmap$ a call map and $\codemap$ a code map. Given a state
  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
  testing property} at a given call depth if:

  \item If the machine fail-stops,
    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
    the property is safisfied for the singleton trace $(\mach, \pol)$ at any
    depth up to the size of $\vs$.

  \item If the top variant of the stack (and only the top variant) returns,
    the property is safisfied for the singleton trace $(\mach, \pol)$ at depth
    equal to the size of the stack. \rb{Avoid confusion with return below}

  \item If the machine takes a step
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    that is neither a call nor a return, all
    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
    (these tests are required also for calls and returns, and omitted in those cases),
    and the stepped $(\mach', \pol')$ and $\vs'$ satisfy the eager testing property
    for a given depth $d$ \rb{omitting stack size condition?} and MP-trace $\MPT$,
    then the property is also satisfied for the extension of the trace \rb{notation}
    at depth $d$.

  \item If the machine takes a call step
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    the contour $\contour$ is updated based on the top of the current machine.
    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    (and in addition to the general step conditions),
    the eager integrity test must hold for $\contour$ at the call,
    the eager confidentiality test must hold for all variants at the call,
    including the new variant that is added to the top of the stack,
    and the property recurses on an arbitrary variant at the new contour
    at depth $d$ with trace $\MPT$.

  \item If the machine takes a return step at the top of the variant stack
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    if (in addition to the general step conditions)
    the property recurses at depth $d$ with trace $\MPT$
    after removing the returning top variant from the stack.

  \item If the machine takes a return step that is also a call, the last two
    cases are combined.

By using property-based testing techniques, we can determine with a high degree
of confidence that an implementation of the conservative micropolicy satisfies
this eager testing property. Although we have not proved that the conservative
policy guarantees the stack safety property, the testing property is provably
stronger than the stack safety property \rb{theorem?} and, by this implication,
we can conclude that the conservative policy also enforces stack safety.

% The properties we give are indeed enforceable.

\rb{What is the strongest property enforced by the conservative policy?}
